Think deep for reef resilience

### 2025-08-11 paper/paper_draft.qmd
Slight updates to DHW-tolerance paragraph and ADRIA-platform paragraph

\[**BG: I am having issues with citations not appearing for datasets like GBRMPA ones.**\]

@@

Our findings suggest that reefs in ‘high’ cover clusters do not have greater thermal
tolerance than reefs in lower coral cover clusters. This is in contrast with literature
suggesting corals have the capacity to adapt to increasing climate change pressures
[@humanes2022; @lachs2024], including the investigation by Bairos-Novak et al.
[-@bairos-novak2021] that revealed high heritability of bleaching and survival response
traits, used to inform thermal tolerance trait distributions in this investigation. The lack
of influence of thermal tolerance on reef outcomes is likely due to reefs being impacted by
high levels of thermal stress early in projected timeseries and being unable to build
thermal tolerance to DHW levels greater than 9°C-weeks. **\[TI note: This paragraph needs to
be revisited\]**

@@

**\[TODO - rewrite/expand paragraphs below for flow\]**

ADRIA - CoralBlox has not been calibrated or validated at the GBR-wide scale. Accordingly,
the purpose of this investigation is to assess the relative influence of reef
characteristics on future resilience and highlight possible outcomes under uncertain future
scenarios. We do not aim to provide spatially explicit predictions of future resilience.

Further analyses could build on the findings of this study by analysing the impacts of
prioritising reef depth in restoration intervention decision making. The ADRIA platform is
developed as a modelling tool to incorporate multi-criteria decision analysis (MCDA) and
ecosystem scenario modelling into reef restoration decision support. The platform allows for
the investigation of different reef characteristics to be targetted in decision making and
models the evolving outcomes as a result of interventions at various locations. The
capability of ADRIA-CoralBlox could be further expanded by incorporating the impacts of
waves, cyclones, CoTS and currents into future resilience investigations [@lin2023;
@mcclanahan2005; @grimaldi2023; @grimaldi2022; @liu2024]. Additionally, the investigation
could be enhanced by incorporating ocean acidification projections into analyses of coral
carbonate budgets.

>>>>

@@

Our findings suggest that reefs in ‘high’ cover clusters do not have greater thermal
tolerance than reefs in lower coral cover clusters. This is in contrast with literature
suggesting corals have the capacity to adapt to increasing climate change pressures
[@humanes2022; @lachs2024], including the investigation by Bairos-Novak et al.
[-@bairos-novak2021] that revealed high heritability of bleaching and survival response
traits, used to inform thermal tolerance trait distributions in this investigation. The low
importance of thermal tolerance on cover outcomes may be due to reefs experiencing high
levels of thermal stress early in projected timeseries. High - early timeseries thermal
stress events may prevent the intergenerational adaptation of corals to thermal stress and
prevent thermal tolerance from reaching levels greater than 9°C-weeks. **\[TI note: This
paragraph needs to be revisited\]**

@@

**\[TODO - rewrite/expand paragraphs below for flow\]** 
As ADRIA - CoralBlox has not been
validated at the GBR-wide scale, we aim to assess the relative influence of reef
characteristics on resilience under future change rather than provide spatially explicit
predictions of reef outcomes into the future. Using this approach we suggest that median
reef depth is an important factor for reef resilience under thermal stress. Improvements in
remote sensing approaches for broad scale bathymetry measurement, and understanding of how
depth impacts experienced thermal stress on reefs will further enhance investigations of
depth as a critical factor for reef resilience. Additional analyses could build on the
findings of this study by assessing the impacts of prioritising reef depth in reef
restoration decision making. The ADRIA platform is developed as a modelling tool to
incorporate multi-criteria decision analysis (MCDA) and ecosystem scenario modelling into
reef restoration decision support. This platform allows for different intervention
priorities to be dynamically targetted in decision making, while modelling the evolving
outcomes as a result of interventions. The capability of ADRIA-CoralBlox could be further
expanded by incorporating the impacts of waves, cyclones, CoTS and currents into future
resilience investigations [@lin2023; @mcclanahan2005; @grimaldi2023; @grimaldi2022;
@liu2024]. Additionally, the investigation could be enhanced by incorporating ocean
acidification projections into analyses of coral carbonate budgets [@eyre2018].
